Liveness-Based Garbage Collection for Lazy Languages

### 2015-07-09 paper/paper.tex

@@

\begin{figure}[t!]
  \psset{unit=1mm}
  \begin{pspicture}(90,48)(0,-48)
    %\psframe(90,48)(0,-48)
    \begin{tabular}{@{}c@{}}
      %\psframebox
          {\sf
	\renewcommand{\arraystretch}{1}{
	  \begin{uprogram}
            \UFL\
            \UNL{0} (\DEFINE\ (\length\  \pl)
	    \UNL{1}  (\SIF~(\NULLQ \ \pl) $0$
            (\PRIM\ 1\ (\CAR\  \pl))))
            \UNL{0}
	    \UNL{0}  (\DEFINE\ (\append\  \lista\ \listb)
	    \UNL{1}  (\SIF~(\NULLQ \ \lista)
	    \listb
	    \UNL{2} \hspace*{-0.1cm}(\CONS\ (\CAR\  \lista) (\append\
            (\CDR\  \lista)
            \listb))))
            \UNL{0}
            \UNL{0} (\LET\ \px\
            $\leftarrow$\ (\CONS\ $5$
            (\CONS\ (\CONS\ $6$ \NIL) \NIL) \IN
	    \UNL{1} (\LET\ \py\   $\leftarrow$\  (\CONS\ $3$ \NIL) \IN
	    \UNL{2}
            (\LET\ \pz\  $\leftarrow$\  (\append\ \px\  \py)\ \IN\
            \UNL{3} (\SIF~(\NULLQ~(\CAR~\pz))~$0$~$\pi$:\
            (\length\ \pz))))))
	  \end{uprogram}
      }}
      \\ 
      (a) Example program. \\  \\
%      \includegraphics[width=.45\textwidth]{motiv-example}
      \input{mem-graph} \\ 
      \renewcommand{\arraystretch}{.9}
      \begin{tabular}[t]{p{0.9\columnwidth}}
        (b) Memory graph at $\pi$.  \scalebox{.7}{\TwoCellsAD{a1}{a2}}
        denotes a closure. Thick edges denote live links. Traversal
        stops at edges marked $\times$ during garbage collection.
      \end{tabular}
    \end{tabular} 
  \end{pspicture}
  \vspace*{-3ex}
  \caption{Example Program and its Memory Graph}\label{fig:mot-example}
\end{figure}

@@

{\color {Myblue}Rewrite at the end: 
Section~\ref{sec:defs} introduces the syntax and semantics of the
language used to illustrate our
analysis.
% Section~\ref{sec:operational} describes the operational semantics of
% the language.
Liveness analysis for this language is  described in
Section~\ref{sec:liveness}. This section also has by  a  sketch  of  a
correctness proof  relative  to  a  non-standard
semantics.  Section~\ref{sec:computing}  shows how to  encode liveness
as   finite-state  automata.    Section~\ref{sec:experiments}  reports
experimental  results  and Section~\ref{sec:lgc-always-better}  proves
that a liveness based collector  can never do more garbage collections
than a reachability based collector.
% and Section~\ref{sec:conclusion} concludes the paper.
}
\section{The target language---syntax and semantics}
\label{sec:defs}
%%We let $x$,  $y$, $z$ range over variables,  $f$ over user-functions
%% and $p$ over primitive functions ($\CONS$, $\PRIM$ etc.).
Figure~\ref{fig:lang-syntax} describes the  syntax of our language. It
is a first order language with lazy semantics. Programs are restricted
to        be        in        Administrative        Normal        Form
(ANF)~\cite{chakravarty03perspective}  where all actual  parameters to
functions  are  variables.  While  this  restriction  does not  affect
expressibility,  this form  has  the benefit  of  making explicit  the
creation  of closures through  the $\LET$  construct.  $\LET$s  in our
language  are lazy;  in the  expression $\LET\,\,  x  \leftarrow s\,\,
\IN\,\, e$, $x$ may occur  in $s$. This enables creation of graph-like
structures in  a pure language. The  restriction of \LET\  to a single
definition  is  for ease  of  exposition---generalization to  multiple
definitions  does  not add conceptual  difficulties.  We  further
restrict each variable  in a program to be distinct,  so that no scope
shadowing   occurs---this  simplifies   proofs   of  soundness.    The
$\SRETURN$ and \SIF\ expressions trigger evaluation of closures.

>>>>

\begin{figure}[t!]
  \psset{unit=1mm}
  \begin{pspicture}(90,48)(0,-48)
    %\psframe(90,48)(0,-48)
    \begin{tabular}{@{}c@{}}
      %\psframebox
          {\sf
	\renewcommand{\arraystretch}{1}{
	  \begin{uprogram}
            \UFL\
            \UNL{0} (\DEFINE\ (\length\  \pl)
	    \UNL{1}  (\SIF~(\NULLQ \ \pl) $0$
            (\PRIM\ 1\ (\CAR\  \pl))))
            \UNL{0}
	    \UNL{0}  (\DEFINE\ (\append\  \lista\ \listb)
	    \UNL{1}  (\SIF~(\NULLQ \ \lista)
	    \listb
	    \UNL{2} \hspace*{-0.1cm}(\CONS\ (\CAR\  \lista) (\append\
            (\CDR\  \lista)
            \listb))))
            \UNL{0}
            \UNL{0} (\LET\ \px\
            $\leftarrow$\ (\CONS\ $5$
            (\CONS\ (\CONS\ $6$ \NIL) \NIL) \IN
	    \UNL{1} (\LET\ \py\   $\leftarrow$\  (\CONS\ $3$ \NIL) \IN
	    \UNL{2}
            (\LET\ \pz\  $\leftarrow$\  (\append\ \px\  \py)\ \IN\
            \UNL{3} (\SIF~(\NULLQ~(\CAR~(\CDR~\pz)))~$0$~$\pi$:\
            (\length\ \pz))))))
	  \end{uprogram}
      }}
      \\ 
      (a) Example program. \\  \\
%      \includegraphics[width=.45\textwidth]{motiv-example}
      \input{mem-graph} \\ 
      \renewcommand{\arraystretch}{.9}
      \begin{tabular}[t]{l}
        (b) Memory graph at $\pi$. Thick edges denote live
        links. Traversal \\{\white (b)} stops at edges
        marked $\times$ during garbage collection.
      \end{tabular}
    \end{tabular} 
  \end{pspicture}
  \vspace*{-3ex}
  \caption{Example Program and its Memory Graph}\label{fig:mot-example}
\end{figure}

@@

@@

Section~\ref{sec:defs} introduces the syntax and semantics of the
language used to illustrate our
analysis.
% Section~\ref{sec:operational} describes the operational semantics of
% the language.
Liveness analysis for this language is  described in
Section~\ref{sec:liveness}. This section also has by  a  sketch  of  a
correctness proof  relative  to  a  non-standard
semantics.  Section~\ref{sec:computing}  shows how to  encode liveness
as   finite-state  automata.    Section~\ref{sec:experiments}  reports
experimental  results  and Section~\ref{sec:lgc-always-better}  proves
that a liveness based collector  can never do more garbage collections
than a reachability based collector.
% and Section~\ref{sec:conclusion} concludes the paper.

\section{The target language---syntax and semantics}
\label{sec:defs}
%%We let $x$, $y$, $z$ range over variables, $f$ over user-functions
and
%% $p$    over    primitive    functions   ($\CONS$,    $\PRIM$ etc.).
Figure~\ref{fig:lang-syntax} describes the  syntax of our language. It
is a first order language with lazy semantics. Programs are restricted
to        be        in        Administrative        Normal        Form
(ANF)~\cite{chakravarty03perspective}  where all actual  parameters to
functions  are  variables.  While  this  restriction  does not  affect
expressibility,  this form  has  the benefit  of  making explicit  the
creation  of closures through  the $\LET$  construct.  $\LET$s  in our
language  are lazy;  in the  expression $\LET\,\,  x  \leftarrow s\,\,
\IN\,\, e$, $x$ may occur  in $s$. This enables creation of graph-like
structures in  a pure language. The  restriction of \LET\  to a single
definition  is  for ease  of  exposition---generalization to  multiple
definitions  does  not add conceptual  difficulties.  We  further
restrict each variable  in a program to be distinct,  so that no scope
shadowing   occurs---this  simplifies   proofs   of  soundness.    The
$\SRETURN$ and \SIF\ expressions trigger evaluation of closures.
